%iffalse
\let\negmedspace\undefined
\let\negthickspace\undefined
\documentclass[journal,12pt,onecolumn]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{txfonts}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{gensymb}
\usepackage{comment}
\usepackage[breaklinks=true]{hyperref}
\usepackage{tkz-euclide} 
\usepackage{listings}
\usepackage{gvv}
\usepackage{circuitikz}
%\def\inputGnumericTable{}                         
\usepackage[latin1]{inputenc}                                
\usepackage{color}                                            
\usepackage{array}                                            
\usepackage{longtable}                                       
\usepackage{calc}                                             
\usepackage{multirow}                                         
\usepackage{hhline}                                           
\usepackage{ifthen}                                           
\usepackage{lscape}
\usepackage{tabularx}
\usepackage{array}
\usepackage{float}
\usepackage{multicol}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{problem}{Problem}
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{example}{Example}[section]
\newtheorem{definition}[problem]{Definition}
\newcommand{\BEQA}{\begin{eqnarray}}
\newcommand{\EEQA}{\end{eqnarray}}
\newcommand{\define}{\stackrel{\triangle}{=}}
\theoremstyle{remark}
\newtheorem{rem}{Remark}

\begin{document}
\bibliographystyle{IEEEtran}
\title{2015-AE-40-52}
\author{EE24BTECH11003 - Akshara Sarma Chennubhatla}
\maketitle
\begin{enumerate}[start=40]

\item The directional derivative of the field $u\brak{x,y,z} = x^2 - 3yz$ in the direction of the vector $\brak{\vec{i} + \vec{j} - 2\vec{k}}$ at point $\brak{2,-1,4}$ is $\rule{2cm}{0.1pt}$
\hfill{\brak{2015}}

\item The composition of an air-entrained concrete is given below:
\begin{align*}
\text{Water} &: 184 kg/m^3\\
\text{Ordinary Portland Cement}\brak{\text{OPC}} &: 368 kg/m^3\\
\text{Sand} &: 606 kg/m^3\\
\text{Coarse aggregate} &: 1155 kg/m^3\\
\end{align*}
Assume the specific gravity of OPC, sand and course aggregate to be $3.14$, $2.67$ and $2.74$, respectively. TYhe air content is $\rule{2cm}{0.1pt}$ litres/$m^3.$
\hfill{\brak{2015}}

\item A bracket plate connected to a column flange transmits a load of $100$ kN as shown in the following figure. The maximum force for which the bolts should be designed is $\rule{2cm}{0.1pt}$ kN.
\hfill{\brak{2015}}
\begin{center}
\input{figs/fig1}
\end{center}

\item Consider the singly reinforced beam section given below $\brak{\text{left figure}}$. The stress block parameters for the cross-section from IS:$456-2000$ are also given below $\brak{\text{right figure}}$. The moment of resistance for the given section by the limit state method is $\rule{2cm}{0.1pt}$ kN-m.
\hfill{\brak{2015}}
\begin{center}
\input{figs/fig2}
\end{center}

\item For formation of collapse mechanism in the following figure, the minimum value of $P_u$ is $xM_p/L$. $M_p$ and $3M_p$ denote the plastic moment capacities of beam section as shown in this figure. The value of $c$ is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2015}}
\begin{center}
\input{figs/fig3}
\end{center}

\item A tapered circular rod of diameter varying from $20$ mm to $10$ mm is connected to another uniform circular rod of diameter $10$ mm as shown in the figure. Both bars are made of same material with the modulus of elasticity, $E = 2$ x $10^5$ MPa. When subjected to a load $P = 30\pi$ kN, the deflection at point $A$ is $\rule{2cm}{0.1pt}$ mm.
\hfill{\brak{2015}}
\begin{center}
\input{figs/fig4}
\end{center}

\item Two beams are connected by a linear spring as shown in the following figure. For a load $P$ as shown in the figure, the percentage of the applied load $P$ carried by the spring is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2015}}
\begin{center}
\input{figs/fig5}
\end{center}

\item For the $2D$ truss with the applied loads shown below, the strain energy in the member $XY$ is $\rule{2cm}{0.1pt}$ kN-m. For the member $XY$, assume $AE = 30$ kN, where $A$ is cross-section area and $E$ is the modulus of elasticity.
\hfill{\brak{2015}}
\begin{center}
\input{figs/fig6}
\end{center}

\item An earth embankment is to be constructed with compacted codesionless soil. The volume of the embankment is $5000m^3$ and the target dry unit weight is $16.2$ kN/$m^3$. Three nearby sites, see the figure below, have been identified from where the required soil can be transported to the construction site. The void ratios $\brak{e}$ of different sites are shown in the figure. Assume the specific gravity of soil to be $2.7$ for all three sites. If the cost of transportation per km is twice the cost of excavation per $m^3$ of borrow pits, which site would you choose as the most economic solution?$\brak{\text{Use unit weight of water} = 10kN/m^3}$
\hfill{\brak{2015}}
\begin{enumerate}
\item Site $X$
\item Site $Y$
\item Site $Z$
\item Any of the sites
\end{enumerate}
\begin{center}
\input{figs/fig7}
\end{center}

\item A water tank is to be constructed on th soil deposit shown in the figure below. A circular footing of diameter $3$ m and depth of embedment $1$ m has been designed to support the tank. The total vertical load to be taken by the footing is $1500$ kN. Assume the unit weight of water as $10$ kN/$m^3$ and the load dispersion pattern as $2V:1H$. The expected settlement of the tank due to primary consolidation of the clay layer is $\rule{2cm}{0.1pt}$ mm.
\hfill{\brak{2015}}
\begin{center}
\input{figs/fig8}
\end{center}

\item A $20$ m thick clay layer is sandwiched between a silty sand layer and a gravelly sand layer. The layer experiences $30$ mm settlement in $2$ years.\\
Given:\\
$T_v = 
\begin{cases}
\frac{\pi}{4}\brak{\frac{U}{100}}^2 & for U \leq 60\% \\
1.781 - 0.933\log_{10}\brak{100-U} & U > 60\%\\
\end{cases}$\\
where $T_v$ is the time factor and $U$ is the degree of consolidation in $\%$.\\
If the coefficient of consolidation of the layer is $0.003$ $cm^2/s$, the deposit will experience a total of $50$ mm settlement in the next $\rule{2cm}{0.1pt}$ years.
\hfill{\brak{2015}}

\item A non-homogeneous soil deposit consiste of a silt later sandwiched between a fine-sand layer at top and a clat later below. Permeability of the silt layer is $10$ times the permeability of the clay layer and one-tenth of the permeability of the sand layer. Thickness of the silt layer is $2$ times the thickness of the sand layer and two-third of the thickness of the clay layer. The ratio of equivalent horizontal and equivalent vertical permeability of the deposit is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2015}}

\item A square footing, $2$ m x $2$ m, is subjected to an inclined point load, $P$ as shown in the figure below. The water table is located below the base of the footing. Considering one-way eccentricity, the net safe load carrying capacity of the footing for a factor of safety of $3.0$ is $\rule{2cm}{0.1pt}$ kN.\\
The following factors may be used:\\
Bearing capacity factors: $N_q = 33.3, N_\gamma = 37.16;$ Shape factors: $F_{qs} = F_{\gamma s} = 1.314;$ Depth factors: $F_{qd} = F_{\gamma d} = 1.113;$ Inclination factors: $F_{qi} = 0.444, F_{\gamma i} = 0.02$
\hfill{\brak{2015}}
\begin{center}
\input{figs/fig9}
\end{center}


\end{enumerate}
\end{document}
