%iffalse
\let\negmedspace\undefined
\let\negthickspace\undefined
\documentclass[journal,12pt,onecolumn]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{txfonts}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{gensymb}
\usepackage{comment}
\usepackage[breaklinks=true]{hyperref}
\usepackage{tkz-euclide} 
\usepackage{listings}
\usepackage{gvv}
\usepackage{circuitikz}
%\def\inputGnumericTable{}                         
\usepackage[latin1]{inputenc}                                
\usepackage{color}                                            
\usepackage{array}                                            
\usepackage{longtable}                                       
\usepackage{calc}                                             
\usepackage{multirow}                                         
\usepackage{hhline}                                           
\usepackage{ifthen}                                           
\usepackage{lscape}
\usepackage{tabularx}
\usepackage{array}
\usepackage{float}
\usepackage{multicol}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{problem}{Problem}
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{example}{Example}[section]
\newtheorem{definition}[problem]{Definition}
\newcommand{\BEQA}{\begin{eqnarray}}
\newcommand{\EEQA}{\end{eqnarray}}
\newcommand{\define}{\stackrel{\triangle}{=}}
\theoremstyle{remark}
\newtheorem{rem}{Remark}

\begin{document}
\bibliographystyle{IEEEtran}
\title{2016-EE-53-65}
\author{EE24BTECH11003 - Akshara Sarma Chennubhatla}
\maketitle
\begin{enumerate}[start=53]

\item A single-phase thyristor-bridge rectifier is fed from a $230 V, 50$ Hz, single-phase AC mains. If it is delivering a constant DC current of $10$ A, at firing angle of $30^\circ$, then value of the power factor at AC mains is
\hfill{\brak{2016}}
\begin{enumerate}
\item $0.87$
\item $0.9$
\item $0.78$
\item $0.45$
\end{enumerate}

\item The switches $T1$ and $T2$ in Figure $\brak{a}$ are switched in a complementary fashion with sinusoidal pulse width modulation technique. The modulating voltage $v_m(t) = 0.8 \sin\brak{200\pi t}$ V and the triangular carrier voltage $\brak{v_c}$ are as shown in Figure $\brak{b}$. The carrier frequency is $5$ kHz. The peak value of the $100$ Hz component of the load current $\brak{i_L}$, in ampere, is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}
\begin{center}
\input{figs/fig1}
\end{center}

\item The voltage $\brak{v_s}$ across and the current $\brak{i_s}$ through a semiconductor switch during a turn-ON transition are shown in figure. The energy dissipated during the turn-ON transition, in mJ, is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}
\begin{center}
\input{figs/fig2}
\end{center}

\item A single-phase $400$ V, $50$ Hz transformer has an iron loss of $5000$ W at the rated condition. When operated at $200$ V, $25$ Hz, the iron loss is $2000$ W. When operated at $416$ V, $52$ Hz, the value of the hysteresis loss divided by the eddy current loss is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}

\item A DC shunt generator delivers $45$ A at a terminal voltage of $220$ V. The armature and the shunt field resistances are $0.01 \Omega$  and $44 \Omega$ respectively. The stray losses are $375$ W. The percentage efficiency of the DC generator is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}

\item A three-phase, $50$ Hz salient-pole synchronous motor has a per-phase direct-axis reactance $\brak{X_d}$ of $0.8$ pu and a per-phase quadrature-axis reactance $\brak{X_q}$ of $0.6$ pu. Resistance of the machine is negligible. It is drawing full-load current at $0.8$ pf $\brak{\text{leading}}$. When the terminal voltage is $1$ pu, per-phase induced voltage, in pu, is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}

\item A single-phase, $22$ kVA, $2200$ V/ $220$ V, $50$ Hz, distribution transformer is to be connected as an auto-transformer to get an output voltage of $2420$ V. Its maximum kVA rating as an auto-transformer is
\hfill{\brak{2016}}
\begin{enumerate}
\item $0.87$
\item $0.9$
\item $0.78$
\item $0.45$
\end{enumerate}

\item A single-phase full-bridge voltage source inverter $\brak{\text{VSI}}$ is fed from a $300$ V battery. A pulse of $120^\circ$ duration is used to trigger the appropriate devices in each half-cycle. The rms value of the fundamental component of the output voltage, in volts, is
\hfill{\brak{2016}}
\begin{enumerate}
\item $234$
\item $245$
\item $300$
\item $331$
\end{enumerate}

\item A single-phase transmission line has two conductors each of $10$ mm radius. These are fixed at a center-to-center distance of $1$ m in a horizontal plane. This is now converted to a three-phase transmission line by introducing a third conductor of the same radius. This conductor is fixed at an equal distance $D$ from the two single-phase conductors. The three-phase line is fully transposed. The positive sequence inductance per phase of the three-phase system is to be $5\%$ more than that of the inductance per conductor of the single-phase system. The distance $D$, in meters, is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}

\item In the circuit shown below, the supply voltage is $10\sin\brak{1000t}$ volts. The peak value of the steady state current through the $1$ Ω resistor, in amperes, is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}
\begin{center}
\input{figs/fig3}
\end{center}

\item A dc voltage with ripple is given by $v\brak{t} = \sbrak{100 + 10\sin\brak{\omega t} - 5\sin\brak{3\omega t}}$ volts. Measurements of this voltage $v\brak{t}$, made by moving-coil and moving-iron voltmeters, show readings of $V_1$ and $V_2$ respectively. The value of $V_2 − V_1$ , in volts, is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}

\item The circuit below is excited by a sinusoidal source. The value of $R$, in $\ohm$, for which the admittance of the circuit becomes a pure conductance at all frequencies is $\rule{2cm}{0.1pt}.$
\hfill{\brak{2016}}
\begin{center}
\input{figs/fig4}
\end{center}

\item In the circuit shown below, the node voltage $V_A$ is $\rule{2cm}{0.1pt}$ V.
\hfill{\brak{2016}}
\begin{center}
\input{figs/fig5}
\end{center}


\end{enumerate}
\end{document}
